prolfqua&colon; A Comprehensive R-package for Proteomics Differential Expression Analysis

### 2023-01-19 inst/MyArticle/paper.Rmd
improve end

When comparing statistical modeling methods for the DEA, we assessed performance measures such as the number of estimated contrasts, the $pAUC$, and if the $FDR$ is an unbiased estimate of the $FDP$. It is relevant that an analysis pipeline shows good performance in all these categories. The examined models *prolfqua_merged*, *proDa*, and *msqrob2* showed good performance in all these categories. Leveraging these computational experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) fit linear models to protein abundances, iii) do not impute missing observation but statistically, model missingness to estimate parameters, i.e., group differences, iv) explicitly report the model used v) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; and last but not least.

>>>>

When comparing statistical modeling methods for the DEA, we assessed performance measures such as the number of estimated contrasts, the $pAUC$, and if the $FDR$ is an unbiased estimate of the $FDP$. An analysis pipeline must demonstrate strong performance in all of these categories. Among the examined models, *prolfqua_merged*, *proDa*, and *msqrob2* all exhibit good performance in all these categories. Leveraging these computational experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) fit linear models to protein abundances, iii) do not impute missing observation but statistically, model missingness to estimate parameters, i.e., group differences, iv) explicitly report the model used v) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; and last but not least.
